### 2022-12-26 manuscript/title.tex
wrote more

\begin{titlepage}
    \begin{center}
        \vspace*{1cm}

        \textbf{The ecological and genetic drivers of silicon accumulation in cereal crops}

        \vspace{1cm}

        by

        \vspace{1cm}

        \textbf{Isaac Peetoom Heida}

        \vfill

        \textsc{a thesis submitted in partial fulfillment of the requirements for the degree of}

        \vspace{1cm}

        \textsc{master of science}

        in

        \textsc{the faculty of graduate and postdoctoral studies}

        (Plant Science)

        \textsc{the university of british columbia}

        (Vancouver)

        \copyright Isaac Peetoom Heida, 2023
    \end{center}
\end{titlepage}