### 2023-08-23 manuscript/coverpage.tex
I think it is ready!

\thispagestyle{plain}
\setcounter{page}{2}
{
%\setlength{\parskip}{12pt}
%\begin{adjustwidth}{0pt}{0pt}
\vspace{1cm}

\noindent The following individuals certify that they have read, and recommend to the Faculty of Graduate and Postdoctoral Studies for acceptance, the thesis entitled:

\vspace{1cm}

\noindent \uline{Ecological and Big Genetic Drivers of Silicon Accumulation in Cereal Crops\hfill}
%\hfill\rule{\textwidth}{0.4pt}

\vspace{1cm}

\noindent submitted by
%\rule{\textwidth}{0.4pt}
\noindent \uline{Isaac Peetoom Heida\hfill}
\noindent in partial fulfilment of the requirements for

\vspace{1cm}
\noindent the degree of \uline{Master of Science\hfill}

\vspace{1cm}
\noindent in \uline{Plant Science\hfill}

\vspace{1cm}

\noindent \textbf{Examining Committee}:

%\vspace{1cm}

\noindent \uline{Juli Carrillo, Professor, Plant Science, UBC\hfill}

%\rule{\textwidth}{0.4pt}

\noindent \textit{Supervisor}

\vspace{1cm}

\noindent \uline{Gurcharn Singh Brar, Professor, Plant Science, UBC\hfill}

%\rule{\textwidth}{0.4pt}

\noindent \textit{Supervisory Committee Member}

\vspace{1cm}

\noindent \uline{Jean-Thomas Cornelis, Professor, Soil Science, UBC\hfill}

%\rule{\textwidth}{0.4pt}

\noindent \textit{Supervisory Committee Member}
%\end{adjustwidth}
}